Generalized Tractability

### 2023-03-10 Older_versions/GeneralizedTractability(old_version).tex
Rearrange old versions and create a new one

\documentclass[11pt,a4paper]{article}
\usepackage{latexsym,amsfonts,amsmath,graphics,mathtools,amssymb,amsthm,booktabs}
\usepackage{epsfig}
%\usepackage[notref,notcite]{showkeys}
%\usepackage[active]{srcltx}
\usepackage{enumitem}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\usepackage{algorithmic}

%\input{common-macros}
%\input{common-macros2}

\input{FJHDef.tex}

\setlength{\textheight}{24cm} \setlength{\textwidth}{16cm}
\setlength{\hoffset}{-1.3cm} \setlength{\voffset}{-1.8cm}

\newcommand{\supp}{\operatorname{supp}}

\newcommand{\tpmod}[1]{{\;(\operatorname{mod}\;#1)}}

\DeclareMathOperator*{\esssup}{ess\,sup}

@@

@@

\newcommand{\fred}[1]{\begingroup\color{blue}Fred: #1\endgroup}
\newcommand{\peter}[1]{\begingroup\color{purple}Peter: #1\endgroup}
\newcommand{\kachi}[1]{\begingroup\color{ForestGreen}#1\endgroup}

@@

\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

@@

TODO: We need to check the relation of our work to that of Michael Gnewuch and Henryk. We do not need tensor product structure: do they need that assumption? The summation conditions on the eigenvalues might be a significant difference between our results and theirs.

\medskip

TODO: Check in how far our conditions coincide with existing conditions for less general tractability notions (polynomial tractability, etc.).

\medskip

Let $\cf_d$ and $\cg_d$ Hilbert spaces, and $\SOL_d : \cf_d \to \cg_d$ be a linear solution operator with adjoint $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors  
\[
\lambda_{\max}^2 \ge \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, 
\]
This means that the non-negative $\lambda_{i,d}$ are the singular values of the $\SOL_d$. The goal is to find an approximate solution, $\APP_d:\cb_d \times (0,\infty) \to \cg_d$ satisfying 
\[
\norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon,
\]
where $\cb_d$ is the unit ball in $\cf_d$ and $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  In this case, the optimal solution is 
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \SOL(u_i) \ip[\cf_d]{f}{u_i},
\]
and the complexity of our linear problem is
\[
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\}.
\]

Certain kinds of common notions of tractability are 
\begin{equation*}
    \begin{array}{r@{\qquad}c@{\qquad}c}
    & \comp(\varepsilon, d)\\
    \toprule
    \text{strong polynomial algebraic  tractability} & \Theta(\varepsilon^{-p})\\
    \text{polynomial algebraic tractability} & \Theta(\varepsilon^{-p}d^{q})\\
    \text{strong polynomial exponential tractability} &  \Theta\bigl([\log(1 + \varepsilon^{-1})]^p\bigr)\\
    \text{polynomial exponential tractability} & \Theta\bigl([\log(1 + \varepsilon^{-1})]^p  d^{q}\bigr)
    \end{array}
\end{equation*}
Let $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ be positive-valued functions that are strictly increasing in both arguments, e.g., 
\[
(x,p) \mapsto x^{p}, \qquad (x,p) \mapsto [\log(1+x)]^p.
\]

All the examples of tractability above are of the form $\comp(\varepsilon, d) = \Theta \bigl( \phi(\varepsilon^{-1},p) \psi(d,q)\bigr)$.  
Note, we define the $\Theta$ notation as follows
\begin{equation*}
     g(\varepsilon,d) = \Theta\bigl(\phi(\varepsilon^{-1},p) \bigr) \text{ means } 
    \lim_{\varepsilon \to 0} \frac{g(\varepsilon,d)}{\phi(\varepsilon^{-1},p')}
           \begin{cases} < \infty, & p'> p,  \\ 
          =\infty, & p' < p.\end{cases}
\end{equation*}
\begin{multline*}
     g(\varepsilon,d) = \Theta\bigl(\phi(\varepsilon^{-1},p) \psi(d,q) \bigr) \text{ means } \\
    \lim_{\substack{\varepsilon \to 0 \\ d \to \infty}} \frac{g(\varepsilon,d)}{\phi(\varepsilon^{-1},p') \psi(d,q')}
          \begin{cases} < \infty, & p'> p, \  q' > q,  \\ 
          =\infty, & p' < p, \ q' < q.\end{cases}
\end{multline*}

\begin{definition}
If there exist the functions $\phi$ and $\psi$ satisfying the conditions listed above and there exist positive numbers $p$ and $q$ with
\[
\comp(\varepsilon, d)  = \Theta\bigl(\phi(\varepsilon^{-1},p) \psi(d,q) \bigr),
\]
then we call a problem $\phi$-$\psi$-tractable with exponents $p$ and $q$.  
If there exists a function $\phi$ satisfying the conditions above and a positive number $p$ such that 
\begin{equation}\label{eq:def_strong_tract}
\comp(\varepsilon, d)  = \Theta\bigl(\phi(\varepsilon^{-1},p) \bigr)
\end{equation}
then we call a problem strongly $\phi$-tractable with exponent $p$. The infimum $p^*$ of all exponents $p>0$ such that \eqref{eq:def_strong_tract} holds, 
is called the exponent of strong $\phi$-tractability. 
\end{definition}

\section{Main Results}
Suppose that, for any real $\tau_1, \tau_2>0$ and any $p,q>0$, it holds that
\begin{align*}
 (\phi(\varepsilon^{-1},p))^{\tau_1} &\le K_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)\\
  (\psi(d,q))^{\tau_2} &\le L_{q,\tau_2} \psi (d,\tau_2 q)\\
\end{align*}
for all $\varepsilon\in (0,1)$, where $K_{p,\tau_1}>0$ is a constant that may depend on $p$ and $\tau_1$, but is independent of $\varepsilon$, 
and where $L_{q,\tau_2}>0$ is a constant that may depend on $q$ and $\tau_2$, but is independent of $d$.

We first show the following result. 
\begin{theorem}\label{thm_main_strong_tract} 
A problem is strongly $\phi$-tractable iff there exists $p>0$ and an integer $C(p)\ge 0$ such that
\begin{equation} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = C(p)}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff} holds, then the exponent of strong $\phi$-tractability equals
\[
 p^* := \inf\{p>0\ \mbox{such that \eqref{eq:tractiff} holds}\}.
\]

\end{theorem}

\begin{proof}
We first show the sufficiency of \eqref{eq:strong_tractiff}. Suppose that \eqref{eq:strong_tractiff} holds for some $p>0$.  
Because $\phi$ is increasing in its arguments, it follows that the complexity may be equivalently expressed as 
\begin{align*}
    \comp(\varepsilon, d) &= \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
    &= \min \biggl\{n \in \natzero : \frac{1}{\phi(\lambda_{n+1, d}^{-1},p)} \le \frac{1}{\phi(\varepsilon^{-1},p)} \biggr\}.
\end{align*}
Since the $\lambda_{i,d}$ are non-increasing it follows that the $\phi(\lambda_{i, d}^{-1},p)$ are non-decreasing. 
In particular, for all $n \in \naturals$,
\begin{align*}
    \MoveEqLeft{\lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{1,d}} \\
    & \implies \frac{1}{\phi(\lambda_{n+1, d}^{-1},p)} \le \frac{1}{\phi(\lambda_{n, d}^{-1},p)} \le \cdots \le \frac{1}{\phi(\lambda_{1, d}^{-1},p)} \\
    & \implies \frac{1}{\phi(\lambda_{n+1, d}^{-1},p)} 
    \le \frac 1n \sum_{i=1}^n  \frac{1}{\phi(\lambda_{i, d}^{-1},p)} 
    \le \frac 1n \sum_{i=1}^\infty  \frac{1}{\phi(\lambda_{i, d}^{-1},p)}.
\end{align*}
Thus, we can conclude that 
\begin{align*}
    n \ge \phi(\varepsilon^{-1},p) \sum_{i=1}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)}
    & \implies 
  \frac 1n \sum_{i=1}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} \le  \frac{1}{\phi(\varepsilon^{-1},p')} \\
   & \implies   \frac{1}{\phi(\lambda_{n+1, d}^{-1},p)} \le \frac{1}{\phi(\varepsilon^{-1},p)}.
\end{align*}
This implies an upper bound on the complexity of
\begin{align*}
       \MoveEqLeft{\comp(\varepsilon,d)} \\
       &\le 1 + \phi(\varepsilon^{-1},p) \sum_{i=1}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} \\
       & = 1 + \phi(\varepsilon^{-1}, p) \left [ \sum_{i=1}^{C(p)-1} \frac{1}{\phi(\lambda_{i, d}^{-1},p)}  
       + \underbrace{\sum_{i=C(p)}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)}}_{=:S_p <\infty\ \text{by \eqref{eq:tractiff}} }\right] \\
       & \le 1+ \phi(\varepsilon^{-1},p) \underbrace{\left(\frac{C(p)-1}{\phi(\lambda_{\max}^{-1},p)} + S_p\right)}_{=:C_1(p)}\\
       & = 1+ \phi(\varepsilon^{-1},p) C_1 (p),
\end{align*}
which means that we have strong $\phi$-tractability, and verifies sufficiency of \eqref{eq:strong_tractiff}.

@@

For the exponent $p^*$ of $\phi$-tractability, we then obviously obtain 
\[
 p^* \le \inf\{p>0\ \mbox{such that \eqref{eq:strong_tractiff} holds}\}.
\]

@@

Let us now show the necessity of \eqref{eq:strong_tractiff}. Indeed, suppose that we have strong 
$\phi$-tractability. I.e., 
\[
\comp(\varepsilon,d)=\Theta (\phi(\varepsilon^{-1},p))
\]
for some $p>0$. This, in turn, means that there is a positive constant $K$ such that
\begin{equation}\label{eq_strong_tract_p'}
\comp(\varepsilon,d)\le K \phi(\varepsilon^{-1},p')
\end{equation}
for all $p'>p$.

Hence, for (possibly small) $\delta_p>0$, there exists a constant $K$ such that \eqref{eq_strong_tract_p'} holds for $p'=p+\delta_p$.

We know that
\[
\lambda_{\comp(\varepsilon,d)+1,d}\le \varepsilon.
\]
Since the eigenvalues $\lambda_{j,d}$ are non-increasing, we have
\begin{equation}\label{eq:lambda_K_strong3}
\lambda_{\lfloor K \phi(\varepsilon^{-1},p')\rfloor +1,d}\le \varepsilon.
\end{equation}
Let
\[
j=j (\varepsilon):= \lfloor K \phi(\varepsilon^{-1},p')\rfloor +1.
\]
If we vary $\varepsilon\in (0,1]$, we see that $j=j_d^*, j_d^*+1, j_d^*+2,\ldots$, where 
\[
  j_d^*=\lfloor K \phi(1,p')\rfloor +1
\]
Note that
\[
j\le K \phi(\varepsilon^{-1},p')+1.
\]
Due to our assumptions, we know that 
$\phi(\cdot, p')$ is increasing in the first argument if 
we keep $p'$ fixed. Let $\phi_{p'}^{-1} (\cdot)$ be the inverse function of $\phi (\cdot, p')$ with respect to the first argument. 
Note that also $\phi_{p'}^{-1}$ is strictly increasing.

Then, 
\[
j \le K \phi(\varepsilon^{-1},p')+1
\]
is equivalent to
\[
\frac{j-1}{K} \le \phi(\varepsilon^{-1},p').
\]
By the monotonicity of $\phi_{p'}^{-1}$, the latter inequality holds if and only if
\[
  \phi_{p'}^{-1} \left(\frac{j-1}{K} \right)
  \le \varepsilon^{-1},
\]
or, equivalently,
\[
 \varepsilon \le \frac{1}{\phi_{p'}^{-1} \left(\frac{j-1}{K} \right)}.
\]
By using \eqref{eq:lambda_K_strong3}, we derive
\[
 \lambda_{j,d}\le \frac{1}{\phi_{p'}^{-1} \left(\frac{j-1}{K} \right)}
\]
for all $j\ge j_d^*$, which is again equivalent to 
\[
\lambda_{j,d}^{-1}\ge \phi_{p'}^{-1} \left(\frac{j-1}{K} \right),
\]
which holds if and only if
\[ 
 \phi (\lambda_{j,d}^{-1}, p') \ge \frac{j-1}{K}.
\]
This implies, for any $\tau>1$, that 
\[ 
 (\phi (\lambda_{j,d}^{-1}, p'))^\tau \ge \frac{(j-1)^\tau}{K^\tau },
\]
which implies, due to our assumption on $\phi$, 
\[ 
 K_{p',\tau}\,\phi (\lambda_{j,d}^{-1}, \tau p') \ge \frac{(j-1)^\tau}{K^\tau }.
\]
This yields
\[
\sum_{j=j_d^*}^\infty \frac{1}{\phi (\lambda_{j,d}^{-1}, \tau p')} \le K_{p',\tau}\, K^\tau \sum_{j=j_d^*}^\infty \frac{1}{(j-1)^\tau}<\infty.
\]

Note, however, that 
\[
 j_d^* = \lfloor K \phi(1,p')\rfloor +1 \le \lfloor K \phi(1,\tau p')\rfloor +1
\]
Hence, we can find $C(\tau p')$ such that 
\[
 \sum_{j=C(\tau p')}^\infty \frac{1}{\phi (\lambda_{j,d}^{-1}, \tau p')}<\infty.
\]
This yields \eqref{eq:strong_tractiff} with $p$ replaced by $\tau p'$, and so we see necessity of $\eqref{eq:strong_tractiff}$.

Assume that $p^*$ is the $\varepsilon$-exponent of strong $\phi$-tractability. We already know from the first part of the proof that 
\[
 p^* \le \inf\{p>0\ \mbox{such that \eqref{eq:strong_tractiff} holds}\}.
\]
Suppose that we would have 
\[
 p^* < \inf\{p>0\ \mbox{such that \eqref{eq:strong_tractiff} holds}\}.
\]
Then, however, we could proceed as in the second part of the proof and show that 
\[
 \sum_{j=C(\tau p')}^\infty \frac{1}{\phi (\lambda_{j,d}^{-1}, \tau p')}<\infty.
\]
for any $\tau>1$, any $p'>p^*$, and a suitable $C(\tau p')$. By choosing $\tau$ sufficiently close to 1, and $p'$ sufficiently close to $p^*$, we then would 
obtain 
\[
 \tau p' < \inf\{p>0\ \mbox{such that \eqref{eq:strong_tractiff} holds}\},
\]
which is a contradiction. So, we must have 
\[
 p^* = \inf\{p>0\ \mbox{such that \eqref{eq:strong_tractiff} holds}\}.
\]

This concludes the proof of the theorem. 
\end{proof}

>>>>

\documentclass{article}
\usepackage{amsmath,amssymb, amsthm, booktabs, mathtools}
\usepackage[utf8]{inputenc}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[notref,notcite]{showkeys}
\input{FJHDef.tex}

@@

%\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

@@

\newcommand{\fred}[1]{\begingroup\color{blue}Fred: #1\endgroup}
\newcommand{\peter}[1]{\begingroup\color{purple}Peter: #1\endgroup}
\newcommand{\kachi}[1]{\begingroup\color{ForestGreen}Kachi: #1\endgroup}

@@

@@

Let $\cf_d$ and $\cg_d$ Hilbert spaces, and $\SOL_d : \cf_d \to \cg_d$ be a linear solution operator with adjoint $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors  
\[
\lambda_{\max}^2 \ge \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots, 
\]
This means that the non-negative $\lambda_{i,d}$ are the singular values of the $\SOL_d$. The goal is to find an approximate solution, $\APP_d:\cb_d \times (0,\infty) \to \cg_d$ satisfying 
\[
\norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon,
\]
where $\cb_d$ is the unit ball in $\cf_d$ and $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  In this case the optimal solution is 
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \SOL(u_i) \ip[\cf_d]{f}{u_i},
\]
and the complexity of our linear problem is
\[
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\}.
\]

Certain kinds of common notions of tractability are 
\begin{equation*}
    \begin{array}{r@{\qquad}c@{\qquad}c}
    & \comp(\varepsilon, d)\\
    \toprule
    \text{strong polynomial algebraic  tractability} & \Theta(\varepsilon^{-p})\\
    \text{polynomial algebraic tractability} & \Theta(\varepsilon^{-p}d^{q})\\
    \text{strong polynomial exponential tractability} &  \Theta\bigl([\log(1 + \varepsilon^{-1})]^p\bigr)\\
    \text{polynomial exponential tractability} & \Theta\bigl([\log(1 + \varepsilon^{-1})]^p  d^{q}\bigr)
    \end{array}
\end{equation*}
Let let $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ be positive-valued functions that are strictly increasing in both arguments, e.g., 
\[
(x,p) \mapsto x^{p}, \qquad (x,p) \mapsto [\log(1+x)]^p.
\]

All the examples of tractability above are of the form $\comp(\varepsilon, d) = \Theta \bigl( \phi(\varepsilon^{-1},p) \psi(d,q)\bigr)$.  Note, we define the $\Theta$ notation as follows
\begin{equation*}
     g(\varepsilon,d) = \Theta\bigl(\phi(\varepsilon^{-1},p) \bigr) \text{ means } 
    \lim_{\varepsilon \to 0} \frac{g(\varepsilon,d)}{\phi(\varepsilon^{-1},p')}
           \begin{cases} < \infty, & p'> p,  \\ 
          =\infty, & p' < p.\end{cases}
\end{equation*}
\begin{multline*}
     g(\varepsilon,d) = \Theta\bigl(\phi(\varepsilon^{-1},p) \psi(d,q) \bigr) \text{ means } \\
    \lim_{\substack{\varepsilon \to 0 \\ d \to \infty}} \frac{g(\varepsilon,d)}{\phi(\varepsilon^{-1},p') \psi(d,q')}
          \begin{cases} < \infty, & p'> p, \  q' > q,  \\ 
          =\infty, & p' < p, \ q' < q.\end{cases}
\end{multline*}
\peter{(PETER: what would happen if $p'<p$ and $q'\ge q$ or the other way around?)}

\begin{definition}
If there exist the functions $\phi$ and $\psi$ satisfying the conditions listed above and there exist positive numbers $p$ and $q$ with
\[
\comp(\varepsilon, d)  = \Theta\bigl(\phi(\varepsilon^{-1},p) \psi(d,q) \bigr),
\]
then we call a problem $\phi$-$\psi$-tractable with exponents $p$ and $q$.  If there exists a function there $\phi$ satisfying the conditions above and a positive number $p$
\[
\comp(\varepsilon, d)  = \Theta\bigl(\phi(\varepsilon^{-1},p) \bigr)
\]
then we call a problem strongly $\phi$-tractable with exponent $p$.
\end{definition}

\section{Main Result}
\begin{theorem}\label{thm_main} 
A problem is $\phi$-$\psi$-tractable with exponents $p$ and $q$ iff there exists a  $C : (0,\infty)^2 \to \naturals$
\begin{equation} \label{eq:tractiff}
    \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = C(p,q)}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} \le 1 \qquad \forall p',q' \in (p,\infty)\times (q,\infty).
\end{equation}
\end{theorem}
\peter{I am wondering whether it would be useful to also include something like "exponents of tractability". I.e., to define minimal $p$ and $q$ such that $\phi-\psi$-tractability holds. I am thinking so because maybe the condition in Theorem 1 should then be modified to hold for all $p'\ge p$ and all $q'\ge q$, and not more generally for all $p,q>0$ as it is now.}

\fred{We may be able to patch up this proof with a condition like $$(\phi(\varepsilon^{-1},p))^{\tau} \le K_{p,\tau} \phi (\varepsilon^{-1},\tau p)$$}
\begin{proof}
Fix $p \in (0,\infty)$.  Because $\phi$ is increasing in its arguments, it follows that the complexity may be equivalently expressed as 
\begin{align*}
    \comp(\varepsilon, d) &= \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
    &= \min \biggl\{n \in \natzero : \frac{1}{\phi(\lambda_{n+1, d}^{-1},p)} \le \frac{1}{\phi(\varepsilon^{-1},p)} \biggr\}.
\end{align*}

Since the $\lambda_{i,d}$ are non-increasing it follows that the $\phi(\lambda_{i, d}^{-1},p)$ are non-decreasing.  In particular, for all $n \in \naturals$,
\begin{align*}
    \MoveEqLeft{\lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{1,d}} \\
    & \implies \frac{1}{\phi(\lambda_{n+1, d}^{-1},p)} \le \frac{1}{\phi(\lambda_{n, d}^{-1},p)} \le \cdots \le \frac{1}{\phi(\lambda_{1, d}^{-1},p)} \\
    & \implies \frac{1}{\phi(\lambda_{n+1, d}^{-1},p)} 
    \le \frac 1n \sum_{i=1}^n  \frac{1}{\phi(\lambda_{i, d}^{-1},p)} 
    \le \frac 1n \sum_{i=1}^\infty  \frac{1}{\phi(\lambda_{i, d}^{-1},p)}.
\end{align*}
Thus, we can conclude that 
\begin{align*}
    n \ge \phi(\varepsilon^{-1},p) \sum_{i=1}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)}
    & \implies 
  \frac 1n \sum_{i=1}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} \le  \frac{1}{\phi(\varepsilon^{-1},p)} \\
   & \implies   \frac{1}{\phi(\lambda_{n+1, d}^{-1},p)} \le \frac{1}{\phi(\varepsilon^{-1},p)}.
\end{align*}
This implies an upper bound on the complexity of
\begin{align*}
       \MoveEqLeft{\comp(\varepsilon,d)} \\
       &\le 1 + \phi(\varepsilon^{-1},p) \sum_{i=1}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} \\
       & = 1 + \phi(\varepsilon^{-1},p) \left [ \sum_{i=1}^{C_3(p,q)-1} \frac{1}{\phi(\lambda_{i, d}^{-1},p)}  + \sum_{i=C_3(p,q)}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} \right ] \\
       & \le 1+  \phi(\varepsilon^{-1},p) \underbrace{\frac{C_3(p,q)-1}{\phi(\lambda_{\max}^{-1},p)}}_{C_1(p,q)}
       +  \phi(\varepsilon^{-1},p) \psi(d,q) 
        \qquad \text{by \eqref{eq:tractiff}} \\
        & \le 1+  \phi(\varepsilon^{-1},p) \bigl[C_1(p,q) + \psi(d,q) \bigr].
\end{align*}

This verifies the sufficient condition.

\peter{PETER: I try to give a proof of the necessary condition (but there is still a gap):

Let us assume that we have $\phi$-$\psi$-tractability. I.e., 
\[
\comp(\varepsilon,d)=\Theta (\phi(\varepsilon^{-1},p),\psi(d,q))
\]
for some $p,q>0$. This, in turn, means that there is a positive constant $K$ such that
\begin{equation}\label{eq_tract_p'q'}
\comp(\varepsilon,d)\le K \phi(\varepsilon^{-1},p')\psi(d,q')
\end{equation}
for all $p'>p$ and $q'>q$.

Hence, for (possibly small) $\delta_p>0$ and $\delta_q>0$, there eists a constant $K$ such that \eqref{eq_tract_p'q'} holds for $p'=p+\delta_p$ 
and $q'=q+\delta_q$.

We know that
\[
\lambda_{\comp(\varepsilon,d)+1,d}\le \varepsilon.
\]
Since the eigenvalues $\lambda_{j,d}$ are non-increasing, we have
\begin{equation}\label{eq:lambda_K}
\lambda_{\lfloor K \phi(\varepsilon^{-1},p'),\psi(d,q')\rfloor +1,d}\le \varepsilon.
\end{equation}
Let
\[
j=j (\varepsilon):= \lfloor K \phi(\varepsilon^{-1},p'),\psi(d,q')\rfloor +1.
\]
If we vary $\varepsilon\in (0,1]$, we see that $j=j_d^*, j_d^*+1, j_d^*+2,\ldots$, where 
\[
  j_d^*=\lfloor K \phi(1,p'),\psi(d,q')\rfloor +1
\]
Note that
\[
j\le K \phi(\varepsilon^{-1},p'),\psi(d,q') +1.
\]
Due to our assumptions, we know that 
$\phi(\cdot, p')$ is increasing in the first argument if 
we keep $p'$ fixed. Let $\phi_{p'}^{-1} (\cdot)$ be the inverse function of $\phi (\cdot, p')$ with respect to the first argument. Note that also $\phi_{p'}^{-1}$ is strictly increasing.

Then, 
\[
j \le K \phi(\varepsilon^{-1},p'),\psi(d,q') +1
\]
is equivalent to
\[
\frac{j-1}{K \psi(d,q')} \le \phi(\varepsilon^{-1},p').
\]
By the monotonicity of $\phi_{p'}^{-1}$, the latter inequality holds if and only if
\[
  \phi_{p'}^{-1} \left(\frac{j-1}{K \psi(d,q')} \right)
  \le \varepsilon^{-1},
\]
or, equivalently,
\[
 \varepsilon \le \frac{1}{\phi_{p'}^{-1} \left(\frac{j-1}{K \psi(d,q')} \right)}.
\]
By using \eqref{eq:lambda_K}, we derive
\[
 \lambda_{j,d}\le \frac{1}{\phi_{p'}^{-1} \left(\frac{j-1}{K \psi(d,q')} \right)}
\]
for all $j\ge j_d^*$, which is again equivalent to 
\[
\lambda_{j,d}^{-1}\ge \phi_{p'}^{-1} \left(\frac{j-1}{K \psi(d,q')} \right),
\]
which holds if and only if
\[ 
 \phi (\lambda_{j,d}^{-1}, p') \ge \frac{j-1}{K \psi(d,q')}.
\]
This implies that 
\[
\frac{1}{\psi(d,q')} \sum_{j=j_d^*}^\infty \frac{1}{\phi (\lambda_{j,d}^{-1}, p')} \le K \sum_{j=j_d^*}^\infty \frac{1}{j-1}.
\]
If the series on the right hand side of the latter inequality would converge (which it obviously does not), then we would be done, since we could argue that we start the summation late enough that the series is less than $1/K$. Then we could argue that this estimate holds for all $d$, and therefore also for the supremum over all $d$. Also the sum of the first $j_d^*$ should be no problem to handle, as this should be bounded by a constant times $\psi (d,q')$.
What we need is a suitable exponent of $j-1$.
}
\end{proof}

@@

\peter{
MODIFIED APPROACH:

@@

\fred{We need to check the proof for 
\begin{itemize}
\item correctness
\item exponents
\item whether the additional assumptions are too restrictive and can be generalized
\item whether we need $\tau$ in the assumption
\end{itemize}}

Let us make the following additional assumptions. Suppose that, for any real $\tau_1, \tau_2>0$ and any $p,q>0$, it holds that
\begin{align*}
 (\phi(\varepsilon^{-1},p))^{\tau_1} &\le K_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)\\
  (\psi(d,q))^{\tau_2} &\le L_{q,\tau_2} \psi (d,\tau_2 q)\\
\end{align*}
for all $\varepsilon\in (0,1)$, where $K_{p,\tau_1}>0$ is a constant that may depend on $p$ and $\tau_1$, but is independent of $\varepsilon$, and where $L_{q,\tau_2}>0$ is a constant that may depend on $q$ and $\tau_2$, but is independent of $d$.

\begin{theorem}\label{thm_main_modified}

A problem is $\phi$-$\psi$-tractable with parameters $p$ and $q$ iff there exists a  $C : (0,\infty)^2 \to \naturals$  such that
\begin{equation} \label{eq:tractiff_modified}
    \sup_{d \in \naturals} \frac{1}{(\psi(d,q'))^\tau} \sum_{i = C(p',q')}^\infty \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau} \le 1 \qquad \forall p',q' \in (p,\infty)\times (q,\infty),\ \forall \tau>1.
\end{equation}
\fred{And the left hand side is infinite if both $p' < p$ and $q' < q$?  Or just one?}
\end{theorem}
\begin{proof}
We first show sufficiency of \eqref{eq:tractiff_modified}. Suppose that \eqref{eq:tractiff_modified} holds for some $p,q>0$. Fix $p'\in (p,\infty)$ and $q'\in (q,\infty)$. Because $\phi$ is increasing in its arguments, it follows that the complexity may be equivalently expressed as 
\begin{align*}
    \comp(\varepsilon, d) &= \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
    &= \min \biggl\{n \in \natzero : \frac{1}{(\phi(\lambda_{n+1, d}^{-1},p'))^\tau} \le \frac{1}{(\phi(\varepsilon^{-1},p'))^\tau} \biggr\}.
\end{align*}

Since the $\lambda_{i,d}$ are non-increasing it follows that the $(\phi(\lambda_{i, d}^{-1},p))^\tau$ are non-decreasing, and this holds for any $\tau>1$.  In particular, for all $n \in \naturals$,
\begin{align*}
    \MoveEqLeft{\lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{1,d}} \\
    & \implies \frac{1}{(\phi(\lambda_{n+1, d}^{-1},p'))^\tau} \le \frac{1}{(\phi(\lambda_{n, d}^{-1},p'))^\tau} \le \cdots \le \frac{1}{(\phi(\lambda_{1, d}^{-1},p'))^\tau} \\
    & \implies \frac{1}{(\phi(\lambda_{n+1, d}^{-1},p'))^\tau} 
    \le \frac 1n \sum_{i=1}^n  \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau} 
    \le \frac 1n \sum_{i=1}^\infty  \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau}.
\end{align*}
Thus, we can conclude that 
\begin{align*}
    n \ge (\phi(\varepsilon^{-1},p'))^\tau \sum_{i=1}^\infty \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau}
    & \implies 
  \frac 1n \sum_{i=1}^\infty \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau} \le  \frac{1}{\phi((\varepsilon^{-1},p'))^\tau} \\
   & \implies   \frac{1}{(\phi(\lambda_{n+1, d}^{-1},p'))^\tau} \le \frac{1}{(\phi(\varepsilon^{-1},p'))^\tau}.
\end{align*}
This implies an upper bound on the complexity of
\begin{align*}
       \MoveEqLeft{\comp(\varepsilon,d)} \\
       &\le 1 + (\phi(\varepsilon^{-1},p'))^\tau \sum_{i=1}^\infty \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau} \\
       &\le 1 +K_{p',\tau} (\phi(\varepsilon^{-1},\tau p')) \sum_{i=1}^\infty \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau} \\
       & = 1 + K_{p',\tau}\phi(\varepsilon^{-1},\tau p') \left [ \sum_{i=1}^{C_3(p',q')-1} \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau}  + \sum_{i=C_3(p',q')}^\infty \frac{1}{(\phi(\lambda_{i, d}^{-1},p'))^\tau} \right ] \\
       & \le 1+ K_{p',\tau} (\phi(\varepsilon^{-1},\tau p')) \underbrace{\frac{C_3(p',q')-1}{(\phi(\lambda_{\max}^{-1},p'))^\tau}}_{C_1(p',q')}
       +  K_{p',\tau}\phi(\varepsilon^{-1},\tau p') (\psi(d,q'))^\tau 
        \qquad \text{by \eqref{eq:tractiff}} \\
        & \le 1+ K_{p',\tau} \phi(\varepsilon^{-1},\tau p') \bigl[C_1(p',q') + L_{q',\tau}\psi(d,\tau q') \bigr].
\end{align*}
As $\tau$ can be arbitrarily close to 1, and $p'$ and $q'$ can be arbitrarily close to $p$ and $q$, respectively, it follows that we have $\phi$-$\psi$-tractability with parameters $p$ and $q$, which shows sufficiency of \eqref{eq:tractiff_modified}.

Let us now show necessity. Indeed, suppose that we have 
$\phi$-$\psi$-tractability with parameters $p$ and $q$. I.e., 
\[
\comp(\varepsilon,d)=\Theta (\phi(\varepsilon^{-1},p),\psi(d,q))
\]
for some $p,q>0$. This, in turn, means that there is a positive constant $K$ such that
\begin{equation}\label{eq_tract_p'q'}
\comp(\varepsilon,d)\le K \phi(\varepsilon^{-1},p'),\psi(d,q')
\end{equation}
for all $p'>p$ and $q'>q$.

Hence, for (possibly small) $\delta_p>0$ and $\delta_q>0$, there eists a constant $K$ such that \eqref{eq_tract_p'q'} holds for $p'=p+\delta_p$ 
and $q'=q+\delta_q$.
